---
title: "solution for q4"
output: html_document
---

<h3><font color='#00008b'>Goal</font></h3>
<p><font size='3'>The goal is to identify any interesting market segments that appear to stand out in NutrientH20's social-media audience.</font></p>
</br>

```{r include=FALSE, echo=FALSE}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(LICORS)
library(knitr)

data = read.csv("social_marketing.csv",row.names=1)
```

<h3><font color='#00008b'>Data Exploration</font></h3>
<p><font size='3'>Some interesting finds from summary of the data:</p>
<p>1. There are 7882 followers in the dataset whose tweets have been categorized into 36 distinct categories.<br />
2. The max number of spam tweets by any follower was 2. This meant that the spam filter was doing its job fairly well.<br /></font></p>

```{r echo=FALSE}
total_col = apply(data,1,sum)
rel_freq = function(x) {
  x / total_col
}

data.1 = data %>% mutate(rel_uncat=rel_freq(uncategorized),rel_adult=rel_freq(adult)) 

data.clean = data.1 %>% filter(rel_uncat<.33&rel_adult<.33)
data.clean = data.clean[,!(names(data.clean) %in% c("rel_uncat","rel_adult","uncategorized","adult","chatter","spam"))]
data.scaled = scale(data.clean,center=TRUE,scale=TRUE)

mu = attr(data.scaled,"scaled:center")
sigma = attr(data.scaled,"scaled:scale")
```

<p><font size='3'>3. On the basis of relative frequency of tweet categories for every followers, we omitted the followers who had more than one-third of their tweets under uncategorized or adult. 26 followers were removed.
4. We will drop uncategorized, adult, chatter and spam columna as they don't add value to cluster definition.<br /></font></p><br/>

<h3><font color='#00008b'>Clustering</font></h3>
<p><font size='3'>First will we perform K-means to find clusters of followers. Number of clusters will be chosen on the basis of elbow plot. The grid for k will be from 2 through 25.</font></p>

```{r echo=FALSE,warning=FALSE,error=FALSE,fig.height=8,fig.width=10,fig.align="center"}
set.seed(1892)
grid=seq(2,25,1)
sse = c()
ch = c()

for(i in grid){
  cluster = kmeans(data.scaled,i,nstart=20)
  sse = c(sse,cluster$tot.withinss)
  W = cluster$tot.withinss
  B = cluster$betweenss
  N = nrow(data.scaled)
  ch1 = (B/W)*((N-i)/(i-1))
  ch = c(ch,ch1)
}

output = as.data.frame(cbind(grid,sse,ch))

ggplot(output) +
  geom_point(aes(x=grid,y=sse)) +
  labs(x='K',y='SSE')
```

<p><font size='3'>In the plot above, we can't see a prominent elbow. We will now plot CH index plot to find value of K.</font></p>

```{r echo=FALSE,fig.height=8,fig.width=10,fig.align="center"}
ggplot(output) +
  geom_point(aes(x=grid,y=ch)) +
  labs(x='K',y='CH')
```

<p><font size='3'>On the basis K vs CH index plot, we will choose k as 6.</font></p>

```{r echo=FALSE}
cluster = kmeans(data.scaled,6,nstart=20)
cluster$size/nrow(data.scaled)
```

<p><font size='3'>The clusters have ~9%, ~8%, ~10%, ~12% and ~64% of the followers.</font></p><br/>

<p><font size='3'>Next we will use K-means ++  initialization.</font><p><br/>

```{r echo=FALSE}
cluster_pp = kmeanspp(data.scaled,6,nstart=20)
cluster_pp$size/nrow(data.scaled)
```

<p><font size='3'>K-means ++ is giving the same clusters sizes. So, we'll stick with K-means result.</font><p><br/>

<p><font size='3'>We now try Hierachical clustering to see if the results change.</font><p><br/>


```{r echo=FALSE}
distance_matrix = dist(data.scaled, method='euclidean')
hier_cluster = hclust(distance_matrix, method='average')
hier_cluster_10 = cutree(hier_cluster, k=10)
summary(factor(hier_cluster_10))
```

<p><font size='3'>Heirachical clustering is not giving good result in this case as one of the cluster is huge with almost 99 percent of the followers.</font><p><br />

<h3><font color='#00008b'>Cluster Analysis</font></h3>
<p><font size='3'>K-means gave us 4 clusters. We will now analyze these clusters.</font><p><br />


```{r echo=FALSE, include=FALSE}
Z = data.clean/rowSums(data.clean)

pc2 = prcomp(Z,scale=TRUE,rank=5,center=TRUE)
summary(pc2)
loadings = pc2$rotation
scores = pc2$x

Z = as.data.frame(cbind(Z,cluster$cluster))
Z$`cluster$cluster` = as.factor(Z$`cluster$cluster`)
names(Z)[names(Z) == 'cluster$cluster'] <- 'cluster'
```

<p><font size='3'>We will use PCA to analyze the cluster. In the plot we see six clusters with position of the followers on the basis on principle component 1 and principle component 2.</font><p><br />

```{r echo=FALSE,fig.height=8,fig.width=10,fig.align="center"}
qplot(scores[,1], scores[,2], color=Z$cluster, xlab='Component 1', ylab='Component 2')+scale_color_colorblind()
```

<br /><p><font size='3'>The order of loadings for principle component 1 is:</font><p>

```{r echo=FALSE} 
colnames(Z)[order(loadings[,1], decreasing=TRUE)]
```

<br /><p><font size='3'>The order of loadings for principle component 2 is:</font><p>

```{r echo=FALSE}
colnames(Z)[order(loadings[,2], decreasing=TRUE)]
```

<br/><p><font size='3'>We will also use biplot to define the clusters.</font><p>

```{r echo=FALSE,fig.height=12,fig.width=12,fig.align="center"}
biplot(pc2)
```


<h3><font color='#00008b'>Conclusion</font></h3>
<p><font size='3'>On the basis of the biplot, principle component 1 and principle component 2 we can define our cluster of followers as:<br/><br/>
1. **Parents/Family oriented**: This cluster has followers who belong baby boomer or gen X category. They tweet a mostly about parenting, food, family, and religion.<br /><br />
2. **Media Junkies**: This cluster has followers who talk about news, politics, business, and currenrt events. They like to talk about movies and television.<br /><br />
3. **College Students**: This cluster has young lads who are in college and like talking about sports and online gaming. They also share photos and love to travel.<br /><br />
4. **Artists**: This cluster has followers who are inclined towards music, crafts, and art.<br/><br />
5. **Fitness Enthusiasts**: This cluster has followers with interests in health and nutrition, personal fitness, and talk about outdoor activities.<br/><br />
6. **Beauty Bloggers**: This cluster has followers who are beauty conscious and talk about fashion and dating.<br/>
</font><p>