---
title: "Portfolio Modeling"
output: html_document
---

```{r include=FALSE, echo=FALSE}
library(mosaic)
library(quantmod)
library(foreach)
```

<br/>
<p><font size='3'>We have designed 3 portfolios keeping in mind the traits of three different type of investors i.e. an agressive investor, a moderate investor, and a conservative investor.<br/><br/>
1. **Agressive Investor**: An agressive investor is a young investor in the age bracket of 18-35, looking for a long term investment of 30 plus years. Such an investor has high tolerance for risk and hence wants a high return.<br/><br/>
2. **Moderate Investor**: A moderate investor is in the age bracket of 35-50. He/she is looking for a mid-term investment horizon and have some risk taking tolerance (not much though). They are looking for moderare returns.<br/><br/>
3. **Safe Investor**: A safe investor is often over the age of 50, nearing retirement. He/she is typically not a risk taker and want their investments to be safe and less volatile. He/she is looking for small but steady returns.</font></p><br/><br/>

<h3><font color='#00008b'>Portfolio for Aggressive Investor</font></h3>
<p><font size='3'>For an agressive investor, we will invest 85% of the wealth in stocks and remaining 15% in bonds. The distribution will be as follows:<br/>
- Large-Cap Stocks (20%): VTI<br/>
- Mid-Cap Stocks (20%): IVOG<br/>
- Small-Cap Stocks (20%): VB<br/>
- International Stocks (15%): SCZ<br/>
- Emerging Markets Stocks (10%): VWO<br/>
- Intermediate Bonds (15%): BND</font></p>

```{r echo=FALSE,include=FALSE,error=FALSE,warning=FALSE}
agg_portfolio = c("VTI","IVOG","VB","SCZ","VWO","BND")

getSymbols(agg_portfolio,from="2014-04-01")

VTIa = adjustOHLC(VTI)
IVOGa = adjustOHLC(IVOG)
VBa = adjustOHLC(VB)
SCZa = adjustOHLC(SCZ)
VWOa = adjustOHLC(VWO)
BNDa = adjustOHLC(BND)

all_returns = cbind(ClCl(VTIa),ClCl(IVOGa),ClCl(VBa),ClCl(SCZa),ClCl(VWOa),ClCl(BNDa))
all_returns = as.matrix(na.omit(all_returns))

N = nrow(all_returns)

initial_wealth = 100000
simulation_agg = foreach(i=1:5000, .combine='rbind') %do% {
  total_wealth = initial_wealth
  weights = c(0.2,0.2,0.2,0.15,0.1,0.15)
  holdings = weights * total_wealth
  n_days = 20
  wealthtracker_agg = rep(0, n_days)
  for(today in 1:n_days) {
    return.today = resample(all_returns, 1, orig.ids=FALSE)
    holdings = holdings + holdings*return.today
    total_wealth = sum(holdings)
    wealthtracker_agg[today] = total_wealth
    holdings = weights * total_wealth
  }
  wealthtracker_agg
}
```

```{r echo=FALSE}
hist(simulation_agg[,n_days]-initial_wealth,breaks=40,main="P&L of Agressive Portfolio",xlab="US dollar",col="blue")
```

```{r include=FALSE}
quantile((simulation_agg[,n_days]- initial_wealth), probs = c(0.05))
```


<p><font size='3'>The Value at Risk of the aggressive portfolio at 5% after running Monte Carlo Simulations is negative USD 4977.34. This means that there is a probablity of 0.05 that the portfolio will fall by USD 4977.34. This can also be interpretted as a loss of USD 4977.34 will be incurred on 1 out of 20 days.</font></p><br/><br />

<h3><font color='#00008b'>Portfolio for Moderate Investor</font></h3>
<p><font size='3'>For a moderate investor, we will invest 70% of the wealth in stocks and remaining 30% in bonds. The distribution will be as follows:<br/>
- Large-Cap Stocks (20%): VTI<br/>
- Mid-Cap Stocks (20%): IVOG<br/>
- Small-Cap Stocks (10%): VB<br/>
- International Stocks (15%): SCZ<br/>
- Emerging Markets Stocks (5%): VWO<br/>
- Intermediate Bonds (30%): BND</font></p>

```{r echo=FALSE,include=FALSE,error=FALSE,warning=FALSE}
mod_portfolio = c("VTI","IVOG","VB","SCZ","VWO","BND")

getSymbols(mod_portfolio,from="2014-04-01")

VTIa = adjustOHLC(VTI)
IVOGa = adjustOHLC(IVOG)
VBa = adjustOHLC(VB)
SCZa = adjustOHLC(SCZ)
VWOa = adjustOHLC(VWO)
BNDa = adjustOHLC(BND)

all_returns = cbind(ClCl(VTIa),ClCl(IVOGa),ClCl(VBa),ClCl(SCZa),ClCl(VWOa),ClCl(BNDa))
all_returns = as.matrix(na.omit(all_returns))

N = nrow(all_returns)

initial_wealth = 100000
simulation_mod = foreach(i=1:5000, .combine='rbind') %do% {
  total_wealth = initial_wealth
  weights = c(0.2,0.2,0.1,0.15,0.05,0.3)
  holdings = weights * total_wealth
  n_days = 20
  wealthtracker_mod = rep(0, n_days)
  for(today in 1:n_days) {
    return.today = resample(all_returns, 1, orig.ids=FALSE)
    holdings = holdings + holdings*return.today
    total_wealth = sum(holdings)
    wealthtracker_mod[today] = total_wealth
    holdings = weights * total_wealth
  }
  wealthtracker_mod
}
```

```{r echo=FALSE}
hist(simulation_mod[,n_days]-initial_wealth,breaks=40,main="P&L of Agressive Portfolio",xlab="US dollar",col="blue")
```

```{r include=FALSE}
quantile((simulation_mod[,n_days]- initial_wealth), probs = c(0.05))
```


<p><font size='3'>The Value at Risk of the moderate portfolio at 5% after running Monte Carlo Simulations is negative USD 3687.38. This means that there is a probablity of 0.05 that the portfolio will fall by USD 3687.38. This can also be interpretted as a loss of USD 3687.38 will be incurred on 1 out of 20 days. As expected, it is lower than that of the aggressive portfolio.</font></p><br/><br />

<h3><font color='#00008b'>Portfolio for Safe Investor</font></h3>
<p><font size='3'>For an safe investor, we will invest 70% of the wealth in stocks and remaining 30% in bonds. The distribution will be as follows:<br/>
- Large-Cap Stocks (25%): VTI<br/>
- Mid-Cap Stocks (10%): IVOG<br/>
- Small-Cap Stocks (10%): VB<br/>
- International Stocks (5%): SCZ<br/>
- Intermediate Bonds (40%): BND<br/ >
- Short-Term Bonds (10%): NEAR</font></p>

```{r echo=FALSE,include=FALSE,error=FALSE,warning=FALSE}
safe_portfolio = c("VTI","IVOG","VB","SCZ","BND","BSV")

getSymbols(safe_portfolio,from="2014-04-01")

VTIa = adjustOHLC(VTI)
IVOGa = adjustOHLC(IVOG)
VBa = adjustOHLC(VB)
SCZa = adjustOHLC(SCZ)
BNDa = adjustOHLC(BND)
BSVa = adjustOHLC(BSV)

all_returns = cbind(ClCl(VTIa),ClCl(IVOGa),ClCl(VBa),ClCl(SCZa),ClCl(BNDa),ClCl(BSVa))
all_returns = as.matrix(na.omit(all_returns))

N = nrow(all_returns)

initial_wealth = 100000
simulation_safe = foreach(i=1:5000, .combine='rbind') %do% {
  total_wealth = initial_wealth
  weights = c(0.25,0.1,0.1,0.05,0.4,0.1)
  holdings = weights * total_wealth
  n_days = 20
  wealthtracker_safe = rep(0, n_days)
  for(today in 1:n_days) {
    return.today = resample(all_returns, 1, orig.ids=FALSE)
    holdings = holdings + holdings*return.today
    total_wealth = sum(holdings)
    wealthtracker_safe[today] = total_wealth
    holdings = weights * total_wealth
  }
  wealthtracker_safe
}
```

```{r echo=FALSE}
hist(simulation_safe[,n_days]-initial_wealth,breaks=40,main="P&L of Agressive Portfolio",xlab="US dollar",col="blue")
```

```{r include=FALSE}
quantile((simulation_safe[,n_days]- initial_wealth), probs = c(0.05))
```


<p><font size='3'>The Value at Risk of the safe portfolio at 5% after running Monte Carlo Simulations is negative USD 2634.098. This means that there is a probablity of 0.05 that the portfolio will fall by USD 2634.098. This can also be interpretted as a loss of USD 2634.098 will be incurred on 1 out of 20 days. As expected, it is the lowest when compared with the aggressive and moderate portfolios.</font></p><br/><br />